---
title: "The truncated tesseract"
author: "Stéphane Laurent"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The truncated tesseract}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

To illustrate the `cxhull` package, we will deal with a four-dimensional 
polytope: the *truncated tesseract*.

It is a convex polytope whose vertices are given by all permutations of

$$
\bigl(\pm 1, \pm(\sqrt{2}+1), \pm(\sqrt{2}+1), \pm(\sqrt{2}+1) \bigr).
$$

Let's enter these 64 vertices in a matrix `points`:

```{r}
sqr2p1 <- sqrt(2) + 1
points <- rbind(
  c(-1, -sqr2p1, -sqr2p1, -sqr2p1),
  c(-1, -sqr2p1, -sqr2p1, sqr2p1),
  c(-1, -sqr2p1, sqr2p1, -sqr2p1),
  c(-1, -sqr2p1, sqr2p1, sqr2p1),
  c(-1, sqr2p1, -sqr2p1, -sqr2p1),
  c(-1, sqr2p1, -sqr2p1, sqr2p1),
  c(-1, sqr2p1, sqr2p1, -sqr2p1),
  c(-1, sqr2p1, sqr2p1, sqr2p1),
  c(1, -sqr2p1, -sqr2p1, -sqr2p1),
  c(1, -sqr2p1, -sqr2p1, sqr2p1),
  c(1, -sqr2p1, sqr2p1, -sqr2p1),
  c(1, -sqr2p1, sqr2p1, sqr2p1),
  c(1, sqr2p1, -sqr2p1, -sqr2p1),
  c(1, sqr2p1, -sqr2p1, sqr2p1),
  c(1, sqr2p1, sqr2p1, -sqr2p1),
  c(1, sqr2p1, sqr2p1, sqr2p1),
  c(-sqr2p1, -1, -sqr2p1, -sqr2p1),
  c(-sqr2p1, -1, -sqr2p1, sqr2p1),
  c(-sqr2p1, -1, sqr2p1, -sqr2p1),
  c(-sqr2p1, -1, sqr2p1, sqr2p1),
  c(-sqr2p1, 1, -sqr2p1, -sqr2p1),
  c(-sqr2p1, 1, -sqr2p1, sqr2p1),
  c(-sqr2p1, 1, sqr2p1, -sqr2p1),
  c(-sqr2p1, 1, sqr2p1, sqr2p1),
  c(sqr2p1, -1, -sqr2p1, -sqr2p1),
  c(sqr2p1, -1, -sqr2p1, sqr2p1),
  c(sqr2p1, -1, sqr2p1, -sqr2p1),
  c(sqr2p1, -1, sqr2p1, sqr2p1),
  c(sqr2p1, 1, -sqr2p1, -sqr2p1),
  c(sqr2p1, 1, -sqr2p1, sqr2p1),
  c(sqr2p1, 1, sqr2p1, -sqr2p1),
  c(sqr2p1, 1, sqr2p1, sqr2p1),
  c(-sqr2p1, -sqr2p1, -1, -sqr2p1),
  c(-sqr2p1, -sqr2p1, -1, sqr2p1),
  c(-sqr2p1, -sqr2p1, 1, -sqr2p1),
  c(-sqr2p1, -sqr2p1, 1, sqr2p1),
  c(-sqr2p1, sqr2p1, -1, -sqr2p1),
  c(-sqr2p1, sqr2p1, -1, sqr2p1),
  c(-sqr2p1, sqr2p1, 1, -sqr2p1),
  c(-sqr2p1, sqr2p1, 1, sqr2p1),
  c(sqr2p1, -sqr2p1, -1, -sqr2p1),
  c(sqr2p1, -sqr2p1, -1, sqr2p1),
  c(sqr2p1, -sqr2p1, 1, -sqr2p1),
  c(sqr2p1, -sqr2p1, 1, sqr2p1),
  c(sqr2p1, sqr2p1, -1, -sqr2p1),
  c(sqr2p1, sqr2p1, -1, sqr2p1),
  c(sqr2p1, sqr2p1, 1, -sqr2p1),
  c(sqr2p1, sqr2p1, 1, sqr2p1),
  c(-sqr2p1, -sqr2p1, -sqr2p1, -1),
  c(-sqr2p1, -sqr2p1, -sqr2p1, 1),
  c(-sqr2p1, -sqr2p1, sqr2p1, -1),
  c(-sqr2p1, -sqr2p1, sqr2p1, 1),
  c(-sqr2p1, sqr2p1, -sqr2p1, -1),
  c(-sqr2p1, sqr2p1, -sqr2p1, 1),
  c(-sqr2p1, sqr2p1, sqr2p1, -1),
  c(-sqr2p1, sqr2p1, sqr2p1, 1),
  c(sqr2p1, -sqr2p1, -sqr2p1, -1),
  c(sqr2p1, -sqr2p1, -sqr2p1, 1),
  c(sqr2p1, -sqr2p1, sqr2p1, -1),
  c(sqr2p1, -sqr2p1, sqr2p1, 1),
  c(sqr2p1, sqr2p1, -sqr2p1, -1),
  c(sqr2p1, sqr2p1, -sqr2p1, 1),
  c(sqr2p1, sqr2p1, sqr2p1, -1),
  c(sqr2p1, sqr2p1, sqr2p1, 1)
)
```

As said before, the truncated tesseract is convex, therefore its convex hull 
is itself. Let's run the `cxhull` function on its vertices:

```{r}
library(cxhull)
hull <- cxhull(points)
str(hull, max = 1)
```

We can observe that `cxhull` has not changed the order of the points:

```{r}
all(names(hull$vertices) == 1:64)
```

Let's look at the cells of the truncated tesseract:

```{r}
table(sapply(hull$facets, function(cell) length(cell$ridges)))
```

We see that 16 cells are made of 4 ridges; these cells are tetrahedra. 
We will draw them later, after projecting the truncated tesseract in the 3D-space.

For now, let's draw the projected vertices and the edges.

The vertices in the 4D-space lie on the centered sphere with radius 
$$
\sqrt{1 + 3 \bigl(\sqrt{2}+1\bigr)^2}.
$$

Therefore, a stereographic projection is appropriate to project the truncated 
tesseract in the 3D-space.

```{r}
sproj <- function(p, r){
  c(p[1], p[2], p[3])/(r-p[4])
}
ppoints <- t(apply(points, 1, 
                   function(point) sproj(point, sqrt(1+3*sqr2p1^2))))
```

Now we are ready to draw the projected vertices and the edges.

```{r, message=FALSE}
edges <- hull$edges
library(rgl)
open3d(windowRect = c(100,100,600,600))
view3d(45,45)
spheres3d(ppoints, radius= 0.07, color = "orange")
for(i in 1:nrow(edges)){
  shade3d(cylinder3d(rbind(ppoints[edges[i,1],],ppoints[edges[i,2],]), 
                     radius = 0.05, sides = 30), col="gold")
}
rglwidget()
```

Pretty nice.

Now let's show the 16 tetrahedra. Their faces correspond to triangular ridges. 
So we get the 64 triangles as follows:

```{r}
ridgeSizes <- sapply(hull$ridges, function(ridge) length(ridge$vertices))
triangles <- t(sapply(hull$ridges[which(ridgeSizes==3)], 
                      function(ridge) ridge$vertices))
head(triangles)
```

Now we repeat the previous plot and we add the triangles:

```{r, message=FALSE}
open3d(windowRect = c(100,100,600,600))
view3d(45,45)
spheres3d(ppoints, radius= 0.07, color = "orange")
for(i in 1:nrow(edges)){
  shade3d(cylinder3d(rbind(ppoints[edges[i,1],],ppoints[edges[i,2],]), 
                     radius = 0.05, sides = 30), col="gold")
}
for(i in 1:nrow(triangles)){
  triangles3d(rbind(
    ppoints[triangles[i,1],],
    ppoints[triangles[i,2],],
    ppoints[triangles[i,3],]),
    color = "red", alpha = 0.4)
}
rglwidget()
```

